---
title: "Rcode: Characteristics of breeding habitat, genetic mating system, and determinants of mating success in the sponge-dwelling goby Elacatinus lori"
author: "Robin Francis"
date: "January 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##datasheet set up
```{r}

#Breeding Habitat Data 

Sponge_size_GROUP_20200311 <- read.csv("C:/Users/robin/Documents/RESEARCH/PAPERS/Determinants of Reproductive Success/MatingSuccess.Rscripts/elori/goby/Sponge_size_GROUP_20200311.csv", na.strings="")
group.data<-Sponge_size_GROUP_20200311
summary(group.data)

#Mating Success Data

datasheet_monthlyAVG_n85 <- read.csv("C:/Users/robin/Documents/RESEARCH/PAPERS/Determinants of Reproductive Success/MatingSuccess.Rscripts/elori/goby/datasheet_monthlyAVG_n85.csv", na.strings="")
data3<-datasheet_monthlyAVG_n85
data3$Month<-as.factor(data3$Month)
str(data3) 
summary(data3)

#for multiple linear regressions (missing some mesurements):

ds3<-na.omit(data3)
summary(ds3)

#Attach packages

library(lme4)
citation("lme4")
library(nlme)
library(lmerTest)
library(MuMIn)
library(sjstats)
library(ggplot2)
library(dplyr)
library(performance)
library(see)
```

##Summary of Sponge Size by Residency Category
```{r}

table(group.data$Group)
prop.table(table(group.data$Group))

summary(group.data$Tube_Length)
parameters::standard_error(group.data$Tube_Length)

summary(group.data$Tube_Length[group.data$Group=="Unoccupied"])
parameters::standard_error(group.data$Tube_Length[group.data$Group=="Unoccupied"])

summary(group.data$Tube_Length[group.data$Group=="Resident"])
parameters::standard_error(group.data$Tube_Length[group.data$Group=="Resident"])

summary(group.data$Tube_Length[group.data$Group=="Male"])
parameters::standard_error(group.data$Tube_Length[group.data$Group=="Male"])
```

##ANOVA of Sponge Size by Residency Category
```{r}

summary(aov(group.data$Tube_Length~group.data$Group))

#or:
group.model<-lm(group.data$Tube_Length~group.data$Group)
group.model.aov<-aov(group.model)
summary(group.model.aov)
group.tukey.test <- TukeyHSD(group.model.aov)
group.tukey.test

#once the t-statistic is above about 30, the p-value maxes out at 2.62e-14

plot(group.tukey.test)
```

##Male Size and Sponge Size 
```{r}

summary(data3$Male_StandardLength_mm)
parameters::standard_error(data3$Male_StandardLength_mm)
hist(data3$Male_StandardLength_mm)
shapiro.test(data3$Male_StandardLength_mm)
#normal dist

summary(data3$Longest_Length_cm)
parameters::standard_error(data3$Longest_Length_cm)
hist(data3$Longest_Length_cm)
shapiro.test(data3$Longest_Length_cm)
#normal dist

#lm of male size ~ sponge size

summary(ds3$Male_StandardLength_mm)
parameters::standard_error(ds3$Male_StandardLength_mm)

summary(ds3$Longest_Length_cm)
parameters::standard_error(ds3$Longest_Length_cm)
  
male_sponge<-lm(ds3$Male_StandardLength_mm~ds3$Longest_Length_cm)

summary(male_sponge)

confint(male_sponge)
performance::r2(male_sponge)
check_autocorrelation(male_sponge)
#ok
```

##Figure 2, Panel a; Frequency of Sponge Sizes by Cat
```{r}

#Breeding Habitat: Sponge Size and Residency Status 

#histograms by residency cat in order to put mult hist on one axis
#NOTE: this plot is NOT stacked, bars are overlayed on top of one another

par(mar=c(5,5,1,1), bty="l")

all.groups_5cm<-hist(group.data$Tube_Length,breaks=seq(1,100,5), bty="l", plot = FALSE)
group.resident<-subset(group.data$Tube_Length, group.data$Group=="Resident")
group.resident_5cm<-hist(group.resident,breaks=seq(1,100,5), bty="l", plot=FALSE)
group.male<-subset(group.data$Tube_Length, group.data$Group=="Male")
group.male_5cm<-hist(group.male,breaks=seq(1,100,5), bty="l", plot=FALSE)

#This is an r plot, could not get the ggplot to look how I wanted, but I tried below


plot(all.groups_5cm,
     bty="l",
     xlim=c(0,100), 
     ylim=c(0,300),
     col = "white",
     main ="", 
     xlab = "Sponge Tube Length (cm)",
     ylab = "Frequency", 
     cex.lab=1.6, 
     cex.axis=1.3)
plot(group.resident_5cm, bty="l", col = "lightgrey", add = TRUE)
plot(group.male_5cm, bty="l", col = "#666666", add = TRUE)
legend("topright",
       legend=c("Unoccupied", "Occupied by Resident", "Occupied by Breeding Male"), 
       fill=c("white", "lightgrey", "#666666"), 
       cex=1.5)
```

##Figure 2 Panel b; Linear Reg Plot of Male Size and Sponge Size
```{r}

par(mar=c(5,5,1,1))
par(bty="n") #this code did not fix
# the panel border/ type of axis (does not match panel a)

ggplot(ds3, aes(Longest_Length_cm, Male_StandardLength_mm))+
  geom_point(shape= 21, color="black", fill="black", size = 1.5)+
  theme_classic()+
  theme(axis.title = element_text(size = 20), 
        axis.title.x = element_text(margin = margin(t = 20)),
        axis.title.y = element_text(margin = margin(t =20, r = 20)),
        axis.text = element_text(size = 18),
        axis.text.x = element_text(colour="black"),
        axis.text.y = element_text(colour="black"))+
  geom_smooth(method="lm", color="black")+
  scale_x_continuous(limits=c(0,100), breaks=c(0,20,40,60,80,100))+
  xlab("Sponge Tube Length (cm)")+
  ylab("Male Standard Length (mm)")
```

##Analyses, Number of Days Paired with a Female 
```{r}
#DAYS PAIRED

summary(ds3$Avg_DaysPaired)
parameters::standard_error(ds3$Avg_DaysPaired)
table(ds3$Avg_DaysPaired)


#clearly right-skewed, should be transformed 
hist(data3$Avg_DaysPaired)
shapiro.test(data3$Avg_DaysPaired)
#not normal

#Log transform 
data3$Log_days<-log(1+data3$Avg_DaysPaired)
ds3$Log_days<-log(1+ds3$Avg_DaysPaired)
hist(data3$Log_days) #looks better? but still not normal
shapiro.test(data3$Log_days) #still not normal


#DAYS PAIRED (LOG TRANSFORMED) MODEL SELECTION

DPL_m1<-lm(Log_days~Male_StandardLength_mm*Longest_Length_cm, data=ds3) 
DPL_m2<-lm(Log_days~Male_StandardLength_mm+Longest_Length_cm, data=ds3)
DPL_m3<-lm(Log_days~Male_StandardLength_mm, data=ds3)
DPL_m4<-lm(Log_days~Longest_Length_cm, data=ds3)
DPL_mNULL<-lm(Log_days~1, data=ds3)

AIC(DPL_m1,DPL_m2,DPL_m3,DPL_m4, DPL_mNULL)
#Lowest = 253 = DPL_m3 or DP ~ Male Size
#within 2 of next lowest AIC so we have to compare models

#smallest model goes first
anova(DPL_m3, DPL_m1)
# m1 is not significantly different from m3
# we chose the simplest model 

#smallest model goes first
anova(DPL_m3, DPL_m2)
# m2 is not significantly different from m3
# we chose the simplest model

anova(DPL_m3, DPL_mNULL)
#m3 is sig diff from null 

### m3 or DP ~ Male Size is the best model 

summary(DPL_m3)
#Male size has a positive effect (for every 1mm in length, male are paired 0.06 more days)

plot(DPL_m3) 
#they look ok
qqplot(ds3$Log_days, ds3$Male_StandardLength_mm)
#looks ok
confint(DPL_m3)
#CI does not cross zero= 95% of the time, 1 mm of male length will have an effect of 0.03-0.09 days paired 
performance::r2(DPL_m3)
#male size expalins 13% of the variation we see in days paired 
check_autocorrelation(DPL_m3)
#ok 
check_model(DPL_m3) 
#looks okay except for normality of residuals 
```

##Figure 3, Days Paired
```{r}
par(mar=c(5,5,1,1))

DaysPaired<-ggplot(ds3, aes(Male_StandardLength_mm, Log_days))+
  geom_point(shape= 21, color="black", fill="black", size = 1.5)+
  theme_classic()+
  theme(axis.title = element_text(size = 20), 
        axis.title.x = element_text(margin = margin(t = 20)),
        axis.title.y = element_text(margin = margin(t =20, r = 20)),
        axis.text = element_text(size = 18),
        axis.text.x = element_text(colour="black"),
        axis.text.y = element_text(colour="black"))+
  geom_smooth(method="lm", 
              color="black")+
  scale_x_continuous(breaks=c(25,30,35,40,45,50,55  ))+
  xlab("Male Standard Length (mm)")+
  ylab(expression(atop("Log Number of Days", paste("Paired with a Female"))))

DaysPaired
```

##Days Paired correlation with other metrics of success
```{r}

#Make a correlation matrix
library("Hmisc")

#need to transform Number of Lays 

#TOTAL LAYS

summary(data3$Avg_TotalLays)
parameters::standard_error(data3$Avg_TotalLays)
hist(data3$Avg_TotalLays)
shapiro.test(data3$Avg_DaysPaired)
#not normal

#Log transform 
data3$Log_Lays<-log(data3$Avg_TotalLays)
ds3$Log_Lays<-log(ds3$Avg_TotalLays)
hist(data3$Log_Lays) #looks better
shapiro.test(data3$Log_Lays) #still not normal, but better

#with data transformations:
dataDP<-select(ds3, Log_days, Log_Lays, Avg_Clutch_Size, Avg_Clutch_Hatch, Avg_Output)
dataDP.cor<-rcorr(as.matrix(dataDP))
dataDP.cor

#inputing results into a table in Word doc for Table 1
```

##Number of Clutches 
```{r}
#TOTAL LAYS

summary(data3$Avg_TotalLays)
parameters::standard_error(data3$Avg_TotalLays)
hist(data3$Avg_TotalLays)
shapiro.test(data3$Avg_DaysPaired)
#not normal

#so we log transformed it to do model selection 

#Log transform 
data3$Log_Lays<-log(1+data3$Avg_TotalLays)
ds3$Log_Lays<-log(1+ds3$Avg_TotalLays)
hist(data3$Log_Lays) #looks better
shapiro.test(data3$Log_Lays) #normal

#"TLL_m" = "Total Lays Log-transformed Model..."

TLL_m1<-lm(Log_Lays~Male_StandardLength_mm*Longest_Length_cm, data=ds3) 
TLL_m2<-lm(Log_Lays~Male_StandardLength_mm+Longest_Length_cm, data=ds3)
TLL_m3<-lm(Log_Lays~Male_StandardLength_mm, data=ds3)
TLL_m4<-lm(Log_Lays~Longest_Length_cm, data=ds3)
TLL_mNULL<-lm(Log_Lays~1, data=ds3)

AIC(TLL_m1,TLL_m2,TLL_m3,TLL_m4, TLL_mNULL)

#Lowest AIC 92, m2
#but with 2 so we need to compare

#smallest model goes first
anova(TLL_m2, TLL_m1)
#m1 not sig diff from m2
#we choose the simpler model 


summary(TLL_m2)
#both male size and sponge size are sig 

### m2 or TLL ~ Male Size AND Sponge Size is the best model 

plot(TLL_m2) 
#they look okay?
confint(TLL_m2)
#CI, 95% conf that 1mm in male size increases total lays by 0.002-0.03
#CI crosses zero for sponge= NOT sig
performance::r2(TLL_m2)
#male and sponge size explains 8% of the variation in total lays
check_autocorrelation(TLL_m2)
#ok 
check_model(TLL_m2) 
#looks ok

```

##Figure 4; Number of clutches
```{r}
#to plot multiple variable models: 

#load car package
library(car)

#Reminder: TLL_m2<-lm(Log_Lays~Male_StandardLength_mm+Longest_Length_cm, data=ds3)

#produce added variable plots
avPlots(TLL_m2)
#NOT FINAL PLOT (see below)

#Here is how to interpret each plot:

#The x-axis displays a single predictor variable, 
  #and the y-axis displays the response variable.
#The blue line shows the association between the predictor variable and the response variable, 
  #while holding the value of all other predictor variables constant.

#The points that are labelled in each plot represent the 2 observations with the largest residuals, 
  #and the 2 observations with the largest partial leverage.

```

##Function to customize Figure 4 
```{r}

#code to make function "AvPlots2" just so that I can change xlabs:

avPlots2 <- function(model, terms=~., intercept=FALSE, layout=NULL, ask, 
                     main, xlab, ...){
    terms <- if(is.character(terms)) paste("~",terms) else terms
    vform <- update(formula(model),terms)
    if(any(is.na(match(all.vars(vform), all.vars(formula(model))))))
        stop("Only predictors in the formula can be plotted.")
    terms.model <- attr(attr(model.frame(model), "terms"), "term.labels")
    terms.vform <- attr(terms(vform), "term.labels")
    terms.used <- match(terms.vform, terms.model)
    mm <- model.matrix(model) 
    model.names <- attributes(mm)$dimnames[[2]]
    model.assign <- attributes(mm)$assign
    good <- model.names[!is.na(match(model.assign, terms.used))]
    if (intercept) good <- c("(Intercept)", good)
    nt <- length(good)
    if (nt == 0) stop("No plots specified")
    if (missing(main)) main <- if (nt == 1) paste("Added-Variable Plot:", good) else "Added-Variable Plots"
    if (nt == 0) stop("No plots specified")
    if (nt > 1 & (is.null(layout) || is.numeric(layout))) {
        if(is.null(layout)){
            layout <- switch(min(nt, 9), c(1, 1), c(1, 2), c(2, 2), c(2, 2), 
                             c(3, 2), c(3, 2), c(3, 3), c(3, 3), c(3, 3))
        }
        ask <- if(missing(ask) || is.null(ask)) prod(layout)<nt else ask
        op <- par(mfrow=layout, ask=ask, no.readonly=TRUE, 
                  oma=c(0, 0, 1.5, 0), mar=c(5, 4, 1, 2) + .1)
        on.exit(par(op))
    }
    if (missing(xlab)) xlab <- paste(good, "| others")
    if (length(xlab) == 1L) xlab <- rep(xlab, length(good))
    if (length(xlab) > length(good))
      warning("'xlab' not length 1 or the number of model names, truncating")
    res <- as.list(NULL)
    for (i in seq_along(good)) {
      term <- good[[i]]
      res[[term]] <- avPlot(model, term, main="", xlab=xlab[[i]], ...)
    }
    mtext(side=3,outer=TRUE,main, cex=1.2)
    invisible(res)
}
```

##Customizing avPlot ; Using this for Figure 4
```{r}

avPlots2(TLL_m2, 
        id=FALSE, 
        col = "black", 
        col.lines = "black", 
        pch = 20, 
        lwd = 2, 
        cex = 1, 
        cex.lab = 1.5, 
        cex.axis = 1.5, 
        xlab =c('Male Standard Length (mm)', 'Sponge Tube Length (cm)'),
        ylab = 'Log Number of Clutches',
        main =" ",
        grid=FALSE,
        ellipse=FALSE,
        marginal.scale=FALSE)


```

##Size of Clutches 
```{r}

hist(data3$Avg_Clutch_Size)
summary(data3$Avg_Clutch_Size)
parameters::standard_error(data3$Avg_Clutch_Size)
table(data3$Avg_Clutch_Size)
shapiro.test(data3$Avg_Clutch_Size)
#normal

#"CS_m" = "Clutch Size Model"
#dataset is the "ds3" (missing values removed) so models are fitted to the same number of observations

CS_m1<-lm(Avg_Clutch_Size~Male_StandardLength_mm*Longest_Length_cm, data=ds3) 
CS_m2<-lm(Avg_Clutch_Size~Male_StandardLength_mm+Longest_Length_cm,  data=ds3)
CS_m3<-lm(Avg_Clutch_Size~Male_StandardLength_mm,  data=ds3)
CS_m4<-lm(Avg_Clutch_Size~Longest_Length_cm,  data=ds3)
CS_mNULL<-lm(Avg_Clutch_Size~1,  data=ds3)

AIC(CS_m1,CS_m2,CS_m3,CS_m4, CS_mNULL)
#Lowest AIC is the NULL model 
```

##Outcome of Clutches 
```{r}

summary(ds3$Avg_Clutch_Hatch)
parameters::standard_error(ds3$Avg_Clutch_Hatch)
table(ds3$Avg_Clutch_Hatch)

summary(data3$Avg_Clutch_Hatch)
parameters::standard_error(data3$Avg_Clutch_Hatch)
hist(data3$Avg_Clutch_Hatch)
shapiro.test(data3$Avg_Clutch_Hatch)
 
#use GLM with quasi-binomial error distr

#"CO_m" = "Clutch Outcome Model..."

CO_m1<-glm(Avg_Clutch_Hatch~Male_StandardLength_mm*Longest_Length_cm, family="quasibinomial", data=ds3) 
CO_m2<-glm(Avg_Clutch_Hatch~Male_StandardLength_mm+Longest_Length_cm, family="quasibinomial", data=ds3)
CO_m3<-glm(Avg_Clutch_Hatch~Male_StandardLength_mm, family="quasibinomial", data=ds3)
CO_m4<-glm(Avg_Clutch_Hatch~Longest_Length_cm, family="quasibinomial", data=ds3)
CO_mNULL<-glm(Avg_Clutch_Hatch~1, family="quasibinomial", data=ds3)

AIC(CO_m1,CO_m2,CO_m3,CO_m4, CO_mNULL)

#will not compute log-lik for quasi models...

#double check models for binomial 
#will get error for using non-integer binomial... 

CO_m1<-glm(Avg_Clutch_Hatch~Male_StandardLength_mm*Longest_Length_cm, family="binomial", data=ds3) 
CO_m2<-glm(Avg_Clutch_Hatch~Male_StandardLength_mm+Longest_Length_cm, family="binomial", data=ds3)
CO_m3<-glm(Avg_Clutch_Hatch~Male_StandardLength_mm, family="binomial", data=ds3)
CO_m4<-glm(Avg_Clutch_Hatch~Longest_Length_cm, family="binomial", data=ds3)
CO_mNULL<-glm(Avg_Clutch_Hatch~1, family="binomial", data=ds3)

AIC(CO_m1,CO_m2,CO_m3,CO_m4, CO_mNULL)
#lowest AIC is null model 


#lets try this: (what we say we do in the methods)

CO2_m1<-lm(Avg_Clutch_Hatch~Male_StandardLength_mm*Longest_Length_cm,  data=ds3) 
CO2_m2<-lm(Avg_Clutch_Hatch~Male_StandardLength_mm+Longest_Length_cm, data=ds3)
CO2_m3<-lm(Avg_Clutch_Hatch~Male_StandardLength_mm,  data=ds3)
CO2_m4<-lm(Avg_Clutch_Hatch~Longest_Length_cm, data=ds3)
CO2_mNULL<-lm(Avg_Clutch_Hatch~1, data=ds3)

AIC(CO2_m1,CO2_m2,CO2_m3,CO2_m4, CO2_mNULL)

#Null is still the best... 

summary(CO2_m3)
#male size not sig 

summary(CO_m4)
#sponge size not sig 
```

##Reproductive Output 
```{r}

summary(data3$Avg_Output)
parameters::standard_error(data3$Avg_Output)
hist(data3$Avg_Output)
shapiro.test(data3$Avg_Output)
#normal

#"RO_m" = "Reproductive Output Model..."

RO_m1<-lm(Avg_Output~Male_StandardLength_mm*Longest_Length_cm, data=ds3) 
RO_m2<-lm(Avg_Output~Male_StandardLength_mm+Longest_Length_cm, data=ds3)
RO_m3<-lm(Avg_Output~Male_StandardLength_mm, data=ds3)
RO_m4<-lm(Avg_Output~Longest_Length_cm, data=ds3)
RO_mNULL<-lm(Avg_Output~1, data=ds3)

AIC(RO_m1,RO_m2,RO_m3,RO_m4, RO_mNULL)

#Lowest AIC 360, m3
#but with 2 so we need to compare

#smallest model goes first
anova(RO_m3, RO_m2)
#m2 not sig diff from m3

summary(RO_m3)
#male size est = 0.07203 p = 0.0245 * 

plot(RO_m3) 
#they look okay?
confint(RO_m3)
#CI, 95% conf that 1mm in male size increases total lays by 0.009-0.13
performance::r2(RO_m3)
#male size explains 5% of the variation in ro
check_autocorrelation(RO_m3)
#ok 
check_model(RO_m3) 
#looks ok
```

##Figure 5, Reproductive Output 
```{r}

DaysPaired<-ggplot(ds3, aes(Male_StandardLength_mm, Avg_Output))+
  geom_point(shape= 21, color="black", fill="black", size = 1.5)+
  theme_classic()+
  theme(plot.margin = margin(2, .1, .1, .1, "cm"),
        axis.title = element_text(size = 20), 
        axis.title.x = element_text(margin = margin(t = 20)),
        axis.title.y = element_text(margin = margin(t =40, r = 20)),
        axis.text = element_text(size = 18),
        axis.text.x = element_text(colour="black"),
        axis.text.y = element_text(colour="black"))+
  geom_smooth(method="lm", 
              color="black")+
  scale_x_continuous(breaks=c(25,30,35,40,45,50,55  ))+
  xlab("Male Standard Length (mm)")+
  ylab("Estimated Reproductive Output")

DaysPaired

```

